Liveness-Based Garbage Collection for Lazy Languages

### 2015-06-05 paper/paper.tex

\subsection{Modified GC Algorithm}
Since the first algorithm requires a lot of bookkeeping and is not
precise, we have a second version of the algorithm 
which does not use liveness for copying closures. 
\begin{verbatim}
procedure copy (cell)
  IF{celltype is cons}
    IF cell is live &&
         cell is not copied using reachability
       new_cell <-  dup_heapcell(cell) 
       new_car        <-  copy(cell.car)
       new_cell.car   <-  new_car
       new_cdr        <-  copy(cell.cdr)
       new_cell.cdr   <-  new_cdr
    ELSE
       cell <- NULL
    ENDIF
  ELSE
    IF {celltype is closure && cell is not copied}
       new_cell      <-  dup_heapcell(cell)
       new_cell.arg1 <-  copy(cell.arg1)
       new_cell.arg2 <-  copy(cell.arg2)
    ENDIF
  ENDIF
 copy_references_on_print_stack()
\end{verbatim}

In this  algorithm we consult  the liveness automata  only if it  is a
cons cell, for  closures we assume everything under it  to be live and
copy everything that is reachable. Although this might lead to copying
more  cells   it  is   faster  as  there   are  no   liveness  checks.
\comment{Once  experimental  results  are  obtained  compare  the  two
  algorithms  here}   Since  we  mix   both  liveness  based   GC  and
reachability the garbage  collector has to be more  careful and ensure
references  marked non-live  are never  copied using  reachability. We
discuss this issue and its solution in sub-section[]\cred{hanging like
  a sore thumb}.

>>>>

\subsection{Modified GC Algorithm}
Since the first algorithm requires a lot of bookkeeping and is not
precise, we have a second version of the algorithm 
which does not use liveness for copying closures. 
\begin{verbatim}
procedure copy (cell)
  IF{celltype is cons}
    IF cell is live &&
         cell is not copied using reachability
       new_cell <-  dup_heapcell(cell) 
       new_car        <-  copy(cell.car)
       new_cell.car   <-  new_car
       new_cdr        <-  copy(cell.cdr)
       new_cell.cdr   <-  new_cdr
    ELSE
       cell <- NULL
    ENDIF
  ELSE
    IF {celltype is closure && cell is not copied}
       new_cell  <- copy_and_mark_reachable_cells(cell)
    ENDIF
  ENDIF
 copy_references_on_print_stack()
\end{verbatim}

In this  algorithm we consult  the liveness automata  only if it  is a
cons cell, for  closures we assume everything under it  to be live and
copy everything that is reachable. Although this might lead to copying
more  cells   it  is   faster  as  there   are  no   liveness  checks.
\comment{Once  experimental  results  are  obtained  compare  the  two
  algorithms here}  Mixing liveness based GC and  reachability has its
own set of pros and cons which we discuss further.\cred{hanging like a
  sore thumb}.
